Outline

### 2020-11-11 text/main.Rmd
last minor edits before sending out to SJB and JMC for last check

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

@@

Similarly, the species-specific vulnerability of functionally-dominant species heavily affects functioning.
By combining species-level vulnerability scores to fishing and climate change induced corall loss [@Graham2011] with the contributions of each species to each function, we demonstrate that loss of individuals most vulnerable to fishing will have greatest impacts on piscivory, followed by P excretion (Fig. 4).
Conversely, loss of individuals due to coral mortality may disproportionally reduce P excretion, N excretion, and production (Fig. 4).
Fishing and loss of live corals both appear to impact species important for P excretion.
Surprisingly, although fishing pressure can negatively affect large herbivores such as parrot fishes [@Bellwood2011], herbivory is the least vulnerable function.
This may be due to the high variability in unique ecosystem roles within the large pool of herbivorous fish species.
While small herbivores are often abundant and not particularly vulnerable to fishing, larger herbivorous species are often highly targeted and are prone to functional extinction in regions with high fishing pressure [@Bellwood2012].
While herbivores of all body sizes and functional groups are combined in our assessment, their realized contributions to herbivory are known to be strongly complementary, and it is important to recognize the specific crucial functional roles performed by subsets of herbivores (e.g. bioerosion [@Bellwood2011; @Green2009]).
Thus, our results reflect the overall vulnerability of functions, yet this does not consider unique roles played by a small group of species, nested in a broad function such as herbivory.

Conserving biomass, diversity, and ecosystem functioning are important objectives of contemporary conservation initiatives [@Cinner2020].
Even though increased fish biomass enhances functions, the strong trade-offs we reveal between key ecosystem functions introduces an as yet unrecognised challenge for coral reef conservation, where actions to enhance one function may negatively affect another.
For example, marine protected areas, which are one of the primary used conservation approaches for coral reefs [@Graham2020], may provide benefits for herbivory. However, marine protected areas do not protect reefs from the pervasive effects of climate change [@Graham2020], and community shifts towards domination of herbivores may affect functions such as P excretion.
Thus, measuring conservation success with biomass or solely one function (e.g. herbivory) can mask the collapse of other essential functions.
Therefore, it is necessary to gauge the state of reef ecosystems based on multiple, complementary, process-based functions [@Brandl2019front], which may lead to informed decisions to prioritize certain functions based on local needs and stressors.
Finally, while there is a general consensus on the role of diversity in enhancing functioning [@Zavaleta2010; @Hooper2005], we highlight the overarching importance of community structure and the identity of dominant species identity at the local scale.
While, it seems clear that maintaining diversity of fishes appears critical for functioning of reefs, our results suggest that species richness only has minor effects on individual functions at local scales. Rather, it may be dissimilarity between local communtiies that is most critical for maintaining function at the seascape scale, as no species appears to consistently provide high contributions for all functions simultaneously or across all sites it occurs [@Duffy2009; @Zavaleta2010].

@@

![Figure 1: World maps of five key ecosystem functions, biomass-corrected and standardised. Dots indicate the locations that are included in this study. Dot size represents the rank of the biomass-corrected function. Similarly, the color scales show the value of the biomass-corrected function categorically (black = lower 25%, grey = 25-75%, color = >75%). Circles highlight the five locations with the highest value of each biomass-corrected function, respectively.](../output/plots/fig1_multimap.png)

![Figure 2: Fixed effect values of the Bayesian linear regression for each function with community variables. Trophic level, size, and immaturity of the community are represented by the median, and lower and upper 95% quantiles of all individuals inside a community. The log-transformed standing stock biomass was also included as a covariable, but not shown in this figure. All data were standardized to compare across functions and variables (See Extended data table 2 for parameter values on non-stansardised data). Functions were log-transformed before standardisation. Dots represent the average of the effect. Horizontal lines indicate the 95% credible interval, with some credible intervals being too small to visualize. Effects are not shown when their 95% CI intersects with zero. ](../output/plots/fig2_com_slopes.png)

![Figure 3: a) Median family-level contributions to each function, relative to their contribution to biomass. Positive values mean that a family contributes more relative to its contribution to total biomass. Only the 12 most important families are shown here, and families are ordered by their median contribution to biomass. b) Distribution of the degree of dominance of communities per function. A degree of dominance of 0 means each species contributes equally to a function, 1 means a single species performs a function. c) Distributions of frequency of being dominant to each function of all species across all communities. A value of 1 means that a species is always important whenever present. A species is counted as being dominant in a community if the contribution of that species is higher than the scenario in which all species contribute equally (i.e. 1 divided by the species richness). Shaded areas show the distribution of the values. Dots represent the median value, while lines indicate the interquartile range.](../output/plots/fig3_contributions.png)

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(drake)
loadd(fig1)
loadd(fig2)
loadd(fig3)
loadd(fig4)
```

@@

Similarly, the species-specific vulnerability of functionally-dominant species heavily affects functioning.
By combining species-level vulnerability scores to fishing and climate change induced coral loss [@Graham2011] with the contributions of each species to each function, we demonstrate that loss of individuals most vulnerable to fishing will have greatest impacts on piscivory, followed by P excretion (Fig. 4).
Conversely, loss of individuals due to coral mortality may disproportionally reduce P excretion, N excretion, and production (Fig. 4).
Fishing and loss of live corals both appear to impact species important for P excretion.
Surprisingly, although fishing pressure can negatively affect large herbivores such as parrot fishes [@Bellwood2011], herbivory is the least vulnerable function.
This may be due to the high variability in unique ecosystem roles within the large pool of herbivorous fish species.
While small herbivores are often abundant and not particularly vulnerable to fishing, larger herbivorous species are often highly targeted and are prone to functional extinction in regions with high fishing pressure [@Bellwood2012].
While herbivores of all body sizes and functional groups are combined in our assessment, their realized contributions to herbivory are known to be strongly complementary, and it is important to recognize the specific crucial functional roles performed by subsets of herbivores (e.g. bioerosion [@Bellwood2011; @Green2009]).
Thus, our results reflect the overall vulnerability of functions, yet this does not consider unique roles played by a small group of species, nested in a broad function such as herbivory.

Conserving biomass, diversity, and ecosystem functioning are important objectives of contemporary conservation initiatives [@Cinner2020].
Even though increased fish biomass enhances functions, the strong trade-offs we reveal between key ecosystem functions introduces an as yet unrecognized challenge for coral reef conservation, where actions to enhance one function may negatively affect another.
For example, marine protected areas, which are one of the primary used conservation approaches for coral reefs [@Graham2020], may provide benefits for herbivory. However, marine protected areas do not protect reefs from the pervasive effects of climate change [@Graham2020], and community shifts towards domination of herbivores may affect functions such as P excretion.
Thus, measuring conservation success with biomass or solely one function (e.g. herbivory) can mask the collapse of other essential functions.
Therefore, it is necessary to gauge the state of reef ecosystems based on multiple, complementary, process-based functions [@Brandl2019front], which may lead to informed decisions to prioritize certain functions based on local needs and stressors.
Finally, while there is a general consensus on the role of diversity in enhancing functioning [@Zavaleta2010; @Hooper2005], we highlight the overarching importance of community structure and the identity of dominant species identity at the local scale.
While, it seems clear that maintaining diversity of fishes appears critical for functioning of reefs, our results suggest that species richness only has minor effects on individual functions at local scales. Rather, it may be dissimilarity between local communities that is most critical for maintaining function at the seascape scale, as no species appears to consistently provide high contributions for all functions simultaneously or across all sites it occurs [@Duffy2009; @Zavaleta2010].

@@

![Figure 1: World maps of five key ecosystem functions, biomass-corrected and standardized. Dots indicate the locations that are included in this study. Dot size represents the rank of the biomass-corrected function. Similarly, the color scales show the value of the biomass-corrected function categorically (black = lower 25%, grey = 25-75%, color = >75%). Circles highlight the five locations with the highest value of each biomass-corrected function, respectively.](../output/plots/fig1_multimap.png)

![Figure 2: Fixed effect values of the Bayesian linear regression for each function with community variables. Trophic level, size, and immaturity of the community are represented by the median, and lower and upper 95% quantiles of all individuals inside a community. We opted to include the lower and upper quantiles of the 95% CI on top of the median to represent the spread of the distribution of sizes, trophic level, and immaturity across all individuals in communities. The log-transformed standing stock biomass was also included as a covariable, but not shown in this figure. All data were standardized to compare across functions and variables (See Extended data table 2 for parameter values on non-standardized data). Functions were log-transformed before standardization. Dots represent the average of the effect. Horizontal lines indicate the 95% credible interval, with some credible intervals being too small to visualize.](../output/plots/fig2_com_slopes.png)

![Figure 3: a) Median family-level contributions to each function, relative to their contribution to biomass. Positive values mean that a family contributes more relative to its contribution to total biomass. Only the 12 most important families are shown here, and families are ordered by their median contribution to biomass. b) Distribution of the degree of dominance of communities per function. A degree of dominance of 0 means each species contributes equally to a function, 1 means a single species performs a function. c) Distributions of frequency of being dominant to each function of all species across all communities. A value of 1 means that a species is always important whenever present. A species is counted as being dominant in a community if the contribution of that species is higher than the scenario in which all species contribute equally (i.e. 1 divided by the number of species that are present and contribute to the function). Shaded areas show the distribution of the values. Dots represent the median value, while lines indicate the interquartile range.](../output/plots/fig3_contributions.png)

### 2020-11-11 text/methods.Rmd
last minor edits before sending out to SJB and JMC for last check

## 1. Underwater visual census database
\noindent We used a published global database of reef fish abundance and sizes collected along belt transects [@Barneche2019]. This database encompasses 9118 transects across 585 sites (98 localities) in the Central Indo-Pacific, Central Pacific, Eastern Pacific, Western Indian, Eastern Atlantic, Western Atlantic. The database only includes sites with a hard reef bottom at the outer reef slope. Transects were carried out at a constant depth, parallel to the reef crest. We selected the species inside families for which we have body stoichiometric data, that were at least 7cm to minimize the bias related to the identification of small individuals, and finally we discarded rare species, for which less than 20 individuals were ever recorded across all transects. The dataset then included 1110 species that belong to 25 families (Acanthuridae, Balistidae, Bothidae, Chaetodontidae, Cirrhitidae, Fistulariidae, Haemulidae, Holocentridae, Kyphosidae, Labridae, Lethrinidae, Lutjanidae, Monacanthidae, Mugilidae, Mullidae, Ostraciidae, Pempheridae, Pomacanthidae, Pomacentridae, Sciaenidae, Scorpaenidae, Serranidae, Siganidae, Tetraodontidae, Zanclidae).

## 2. Bioenergetic modeling
\noindent We applied a series of bioenergetic models for each existing combination of species identity, body size, and sea surface temperature (n = 30668). The bioenergetic model framework integrates elements of metabolic theory, stoichiometry, and flexible elemental limitation [@Schiettekatte2020]. As such, this model is well suited for the application of a wide variety of fish species. The model relies on a set of input parameters including elements of metabolism, growth, and diet and body stoichiometry. We quantified the input parameters for all species through the integration of empirical data, data synthesis, and Bayesian phylogenetic models (See supplementary methods). The bioenergetic model then allows for the prediction of elemental fluxes mediated by individuals, including ingestion rate, excretion rates of N and P, and growth rate. Here, we focused on 5 key processes: N excretion rate (g N / day), P excretion rate (g P/day), production of body mass through growth (g C/day), herbivory, i.e. ingestion rate of macrophytes (g/day), and piscivory, i.e. ingestion rate of fishes (g/day). We summed the contribution of each individual to each process on the transect-level, to get community-level estimates, then dividing it by the surface area that was surveyed. Each process is thus expressed in dry mass per day per square meter. We note that N excretion, P excretion, and production include contributions of all fishes, whereas herbivory and piscivory are only carried out by a subset of the community, with respect to their trophic guild [@Parravicini2020]. To reduce the occurrence of misclassification of herbivores and piscivores, we only categorized a species as a herbivore or piscivore if it had both the highest probability to be classified in that trophic group and this probability was more than 0.5, based on the probability scores of trophic guilds for a global fish species database [@Parravicini2020]. Further, as a comparison, we quantified herbivory and piscivory rates using two alternative trophic guild classifications based on expert opinion (Extended data figure 5). We can conclude that overall the herbivory and piscivory rates are very similar across trophic guild classifications, we are slightly conservative in piscivory rate estimations, and herbivory rates are almost equal to rates based on Siqueira et al. (2020) [@Siqueira2020], and exceed estimations based on Mouillot et al. (2014) [@Mouillot2014], which can be explained by the fact that this database classified many herbivores as omnivores. Thus, we opted to use the trophic classification by Parravicini et al. (2020), that defines trophic guilds based on empirical data using a quantitative, unbiased, and fully reproducible framework [@Parravicini2020].

@@

## 4. Effect of community structure on ecosystem functions 
\noindent To investigate the effect of the community structure while still accounting for the effects of standing biomass and sea surface temperature, we quantified a set of variables that characterize the community. These variables describe the size, age, and trophic distribution of the community, as these may all affect functions [@Schiettekatte2020]. Specifically, we calculated the 2.5%, 50% and 97.5% quantiles of the total length, immaturity, and trophic level of all individuals per transect. The total length is based on the visual estimation by divers. The immaturity is quantified using the following formula: 
<!-- \begin{equation} -->
	$$\textrm{immaturity}_\textrm{i} = \kappa (\textrm{l}_{\infty} - \textrm{l}_\textrm{i}),$$
<!-- 	\label{imm} -->
<!-- \end{equation} -->
where $\kappa$ is the species-specific growth rate parameter and $l_{\infty}$ is the species-specific asymptotic adult length, and $l_i$ is the total length of individual i. Essentially, this is the derivative of the Von Bertalanffy growth model for a certain length, and the higher this value is, the younger the individual. Finally, the trophic level was extracted from fishbase [@Froese2018]. Additionally, we quantified the transect-level species richness. 
For each log-transformed function we then fitted a Bayesian mixed-effect model with all 12 above-mentioned variables:, after verifying that there are no strong correlations between variables (the highest correlation coefficient was 0.5, and 50% of the variable pair correlations varied between -0.1 and 0.2).

@@

To compare effects across functions and assess the relative importance of each variable, we standardized all variables prior to model fitting. We fitted all 5 models by using 4 cores, that each had 2000 iterations with a warm-up of 1000 iterations. We used weakly-informative priors [@Burkner2017].

## 5. Species dominance and contributions to functions
To investigate the role of species for each function, we quantified the contribution of each species to each function for all transects as followed:
<!-- \begin{equation} -->
$$	\textrm{contribution}_{f,i,j} = \frac{\sum F_{f,i,j}}{\sum F_{f,j}},$$
<!-- 	\label{con} -->
<!-- \end{equation} -->
where i is a certain species, j is a transect, F is the value of function f.

@@

where $C_i$ is the contribution of a certain species and R is the species richness. The degree of dominance thus ranges between 0 and 1, where 0 means that each species contributes equally and 1 means that one species performs the entire function.

>>>>

## 1. Underwater visual census database
\noindent We used a published global database of reef fish abundance and sizes collected along belt transects [@Barneche2019]. This database encompasses 9118 transects across 585 sites (98 localities) in the Central Indo-Pacific, Central Pacific, Eastern Pacific, Western Indian, Eastern Atlantic, Western Atlantic. The database only includes sites at the outer reef slope and with a hard reef bottom. Transects were carried out at a constant depth, parallel to the reef crest. We selected the species inside families for which we have body stoichiometric data, that were at least 7cm to minimize the bias related to the identification of small individuals, and finally we discarded rare species, for which less than 20 individuals were ever recorded across all transects. The dataset then included 1110 species that belong to 25 families (Acanthuridae, Balistidae, Bothidae, Chaetodontidae, Cirrhitidae, Fistulariidae, Haemulidae, Holocentridae, Kyphosidae, Labridae, Lethrinidae, Lutjanidae, Monacanthidae, Mugilidae, Mullidae, Ostraciidae, Pempheridae, Pomacanthidae, Pomacentridae, Sciaenidae, Scorpaenidae, Serranidae, Siganidae, Tetraodontidae, Zanclidae).

## 2. Bioenergetic modeling
\noindent Here, we focused on 5 key processes mediated by fish: N excretion rate (gN/day/m^2^), P excretion rate (gP/day/m^2^), production of body mass through growth (gC/day/m^2^), herbivory, i.e. ingestion rate of macrophytes (gC/day/m^2^), and piscivory, i.e. ingestion rate of fishes (g/day/m^2^). These 5 processes were estimated in each transect using individual-based bioenergetic models that predicts elemental fluxes, including ingestion rate, excretion rates of N and P, and growth rate.
The bioenergetic model framework integrates elements of metabolic theory, stoichiometry, and flexible elemental limitation [@Schiettekatte2020]. We quantified the input parameters, including elements of metabolism, growth, and diet and body stoichiometry, for all 1110 species through the integration of empirical data, data synthesis, and Bayesian phylogenetic models (See supplementary methods). We ran the model for each combination of species identity, body size, and sea surface temperature (n = 30668) to get the contribution of each individual to each process in each transect and the cumulated estimates for the fish community per surface area. Each process is thus expressed in dry mass per day per square meter. We note that N excretion, P excretion, and biomass production include contributions of all fishes, whereas herbivory and piscivory are carried out by a subset of the community, with respect to their trophic guild [@Parravicini2020]. To reduce the occurrence of misclassification of herbivores and piscivores, we categorized a species as a herbivore or piscivore if it had both the highest probability to be classified in that trophic group and this probability was more than 0.5, based on the probability scores of trophic guilds for a global fish species database that defines trophic guilds based on empirical data using a quantitative, unbiased, and fully reproducible framework [@Parravicini2020]. Further, as a comparison, we quantified herbivory and piscivory rates using two alternative trophic guild classifications based on expert opinion [@Mouillot2014; @Parravicini2020] (Extended data figure 5). Both the herbivory and piscivory rates are congruent with the expert opinion trophic guild classifications.

@@

## 4. Effect of community structure on ecosystem functions 
\noindent To investigate the effect of the community structure while still accounting for the effects of standing biomass and sea surface temperature, we quantified a set of variables that characterize the community. These variables describe the size, age, and trophic distribution of the community, as these may all affect functions [@Schiettekatte2020]. Specifically, we calculated the 2.5%, 50% and 97.5% quantiles of the total length, immaturity, and trophic level of all individuals per transect. The total length is based on the visual estimation by divers. The immaturity is quantified using the following formula: 
<!-- \begin{equation} -->
	$$\textrm{immaturity}_\textrm{i} = \kappa (\textrm{l}_{\infty} - \textrm{l}_\textrm{i}),$$
<!-- 	\label{imm} -->
<!-- \end{equation} -->
where $\kappa$ is the species-specific growth rate parameter and $l_{\infty}$ is the species-specific asymptotic adult length, and $l_i$ is the total length of individual i. Essentially, this is the derivative of the Von Bertalanffy growth model for a certain length, and the higher this value is, the younger the individual. Finally, the trophic level was extracted from fishbase [@Froese2018]. Additionally, we quantified the transect-level species richness. 
For each log-transformed function we then fitted a Bayesian mixed-effect model with all 12 above-mentioned variables, after verifying that there are no strong correlations between variables (the highest correlation coefficient was 0.5, and 50% of the variable pair correlations varied between -0.1 and 0.2).

@@

To compare effects across functions and assess the relative importance of each variable, we standardized all variables prior to model fitting. We fitted all 5 models by using 4 cores, that each had 2000 iterations with a warm-up of 1000 iterations, and used weakly-informative priors [@Burkner2017].

## 5. Species dominance and contributions to functions
We quantified the relative contribution of each species to each function for all transects as followed:
<!-- \begin{equation} -->
$$	\textrm{contribution}_{f,i,j} = \frac{\sum F_{f,i,j}}{\sum F_{f,j}},$$
<!-- 	\label{con} -->
<!-- \end{equation} -->
where i is a certain species, j is a transect, F is the value of function f.

@@

where $C_i$ is the contribution of a certain species and R is the number of species contributing to a certain function. The degree of dominance thus ranges between 0 and 1, where 0 means that each species contributes equally and 1 means that a single species performs the entire function. In the case of N excretion, P excretion, and production, R equals the species richness, while for herbivory and piscivory R represents the number of herbivores and piscivores, respectively.

### 2020-11-11 text/suppl_tables_figures.Rmd
last minor edits before sending out to SJB and JMC for last check

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, error = FALSE)
library(tidyverse)
library(flextable)
library(officer)
library(drake)
library(broom)
```

@@

  flextable(data) %>%
    fontsize(size = 8) %>%
    fontsize(size = 10, part = "header") %>%
    set_caption("Extended Date table 2: Overview of localities of UVC transects, used in this study, including number of sites and number of transects") %>%
    autofit()
head(data)
```

@@

![Extended Data Figure 5: Comparison herbivory and piscivory rates when using alternative diet classifications](../output/plots/Supplemental_herbivory_piscivory_alternative_classification.png)

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, error = FALSE)
library(tidyverse)
library(flextable)
library(officer)
library(drake)
library(broom)

loadd(SI_fig1)
loadd(SI_fig2)
loadd(SI_pp_plots)
loadd(SI_rank_plots)
```

@@

  flextable(data) %>%
    fontsize(size = 8) %>%
    fontsize(size = 10, part = "header") %>%
    set_caption("Extended Date table 2: Overview of localities of UVC transects, used in this study, including number of sites and number of transects") %>%
    autofit()
```

@@

![Extended Data Figure 5: Comparison herbivory and piscivory rates when using alternative diet classifications from Mouillot et al. (2014) and Siqueira et al. 2020](../output/plots/Supplemental_herbivory_piscivory_alternative_classification.png)
